Nonstationary, Quasi-Periodic Risk of Climate Extremes, and Decision Analysis

### 2018-02-21 writeup/main.tex
put floodsampling package in here for simplicity

To incorporate secular change into a statistical model, many studies have used time parameterizations on certain parameters.
We consider a trend model where the data is log-normally distributed as \cref{eq:lognormal} with a linear time trend on both \( \mu \) and \( \sigma \):
\begin{align}
  \log Q_\text{hist} & \sim \normal \qty(\mu_0 + \beta_\mu \qty(t - t_0), \sigma_0 + \beta_\sigma \qty(t-t_0)) \\
  \mu_0 & \sim \normal \qty(0, 10) \\
  \beta_\mu & \sim \normal \qty(0, 0.1) \\
  \sigma_0 & \sim \normal \qty(0, 2) \\
  \beta_\sigma & \sim \normal \qty(0, 0.05)
\end{align}
As for the stationary case, we fit posterior samples in a Bayesian framework using \texttt{stan} and apply weak priors.
The influence of priors in this case is stronger than in the stationary case, and a reasonable choice of priors might more strongly favor \(\beta_mu=0\) and \(\beta_\sigma=0\), for example using the Horseshoe prior parameterization \citep{Piironen2016a}.
However, in the context of a synthetic data exercise it seems reasonable to apply priors that are at most weakly informative.

@@

For a given choice of \(M\), \(N\), generating model, and fitting model, \(J\) sequences of streamflow of length \(M+N\) years each are generated.
For each of the \(J\) sequences, the fitting model is then fit on the \(N\)-year historical record.
\(K\) posterior simulations of future rainfall are then drawn from the fitted model.
We then perform a standard decomposition of uncertainty:
\begin{equation}
  \mathbb{E} \qty[ \qty( \hat{p}_T - p_T )^2 ] = \qty( \mathbb{E} \qty[ \hat{p}_T ] - p_T )^2 + \mathbb{E} \qty[ \qty( \hat{p}_T - \mathbb{E} \qty[ \hat{p}_T ] )^2]
\end{equation}
where the first term is the mean squared error, the second term is the bias squared, and the third term is the variance.

@@

\section{Results}

@@

>>>>

To incorporate secular change into a statistical model, many studies have used time parameterizations on certain parameters.
We consider a trend model where the data is log-normally distributed as \cref{eq:lognormal} with a linear time trend on both \( \mu \) and \( \sigma \).
As for the stationary case, we fit posterior samples in a Bayesian framework using \texttt{stan} and apply weak priors:
\begin{align}
  \log Q_\text{hist} & \sim \normal \qty(\mu_0 + \beta_\mu \qty(t - t_0), \sigma_0 + \beta_\sigma \qty(t-t_0)) \\
  \mu_0 & \sim \normal \qty(0, 10) \\
  \beta_\mu & \sim \normal \qty(0, 0.1) \\
  \sigma_0 & \sim \normal \qty(0, 2) \\
  \beta_\sigma & \sim \normal \qty(0, 0.05)
\end{align}
The influence of priors in this case is stronger than in the stationary case, and an alternative choice of priors might more strongly favor \(\beta_\mu=0\) and \(\beta_\sigma=0\), for example using the Horseshoe prior parameterization \citep{Piironen2016a}.

@@

For a given choice of \(M\), \(N\), generating model, and fitting model, \(J\) sequences of streamflow of length \(M+N\) years each are generated.
For each of the \(J\) sequences, the fitting model is then fit on the \(N\)-year historical record.
\(K\) posterior simulations of future rainfall are then drawn from the fitted model.
We then perform a standard decomposition of uncertainty:
\begin{equation}
  \mathbb{E} \qty[ \qty( \hat{p}_T - p_T )^2 ] = \qty( \mathbb{E} \qty[ \hat{p}_T ] - p_T )^2 + \mathbb{E} \qty[ \qty( \hat{p}_T - \mathbb{E} \qty[ \hat{p}_T ] )^2]
\end{equation}
where the first term is the mean squared error, the second term is the bias squared, and the third term is the variance.
In \cref{sec:results} the bias and standard deviation (square root of variance) are averaged over all \(J\) sequences and plotted.\footnote{Check this code}

@@

\section{Estimation Bias and Uncertainty}\label{sec:results}

@@

\emph{Add more text once the results come through from Habanero -- they are running now!}
